An algorithm for finding the global extremum of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-08-29 2023/MMST/samplepaper.tex
change the section 3 header

\section{Solving Method}

@@

\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach \#4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach \#5} \label{fig5}
\end{figure}

@@

        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable point and one boundary point}, apply the original rules of the global search algorithm
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable point and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
\end{enumerate}

@@

\begin{table}[h]
\caption{Comparison of the original and modified GSA}\label{tab1}
\begin{tabular}{|l|c|c|c|}
\hline
 &  classical  &  round district  &  random district  \\
\hline
Average number of trials & 781,66	& 854,91 & 839,03 \\
Average number of trials in non-computability regions & - & 185,79 & 42,31 \\
Number of solved problems & 100 / 100 & 100 / 100 & 100 / 100\\
\hline
\end{tabular}
\end{table}

>>>>

\section{Method for Solving the Problems with Partially Defined Objective Function}

@@

\subsubsection{The fourth approach.} To restore the objective function values in the points for the intervals with two non-computable points as the averages of the nearest computable points taking into account the coordinates, i.e.
\begin{equation}\label{eq17} 
%\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \| x-x' \|}{\| x''-x' \|},
\tilde{z}(x)=z'+ \text{sign}(z''-z') \frac {|z''-z'| \cdot \rho(x,x')}{\rho(x',x'') },
\end{equation}
where $z'$ and $z''$ are the objective function values in the computable points $x'$ and $x''$  nearest to the interval and $\rho(x_1,x_2) =  |x_1 - x_2|^{1/N}$.
The rest of the approach is identical to the third one. Formula (\ref{eq14}) in this case will be applied when the non-computable region is located only near the boundary point.
\begin{figure}
\includegraphics[width=\textwidth]{fig4.png}
\caption{Imputation of non-computable values by averages of the neighboring points. Approach No. 4} \label{fig4}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{fig5.png}
\caption{Imputation of non-computable values by averages of the minorant and the majorant. Approach No. 5} \label{fig5}
\end{figure}

@@

        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable and one boundary point}, apply the original rules of the global search algorithm
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
\end{enumerate}

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{fig6.png}
\caption{Examples of problems No.No. 19, 25, and 35 from the first series of experiments} \label{fig6}
\end{figure}

\begin{table}[h]
\caption{Comparison of the original and modified GSA}\label{tab1}
\centering
\begin{tabular}{|l|c|c|c|}
\hline
 &  { classical }  & { round district }  & { random district }  \\
\hline
Average number of trials & 781.7	& 854.9 & 839.0 \\
\begin{tabular}{@{}l@{}} {Average number of trials } \\  { in non-computability regions } \end{tabular}  & - & 185.8 & 42.3 \\
{Number of solved problems } & 100 / 100 & 100 / 100  &  100 / 100 \\
\hline
\end{tabular}
\end{table}
